---
title: "R_Markdown_Graphs"
author: "Amb"
date: "2023-12-19"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load library, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(vioplot)
library(lattice)
library(ggplot2)
library(thematic)
library(ragg)
```


```{r cars}
# Define the cars vector with 5 values
cars <- c(1, 3, 6, 4, 9)
```

```{r autos_data_df}
autos_data <- data.frame (cars  = c(1, 3, 6, 4, 9),
                  trucks = c(2, 5, 4, 5, 12),
                  suvs = c(4,4,6,6,16)
                )

```

```{r linechart_1_scatterplot}
# Graph the cars vector with all defaults
plot(cars)

```


```{r linechart_2}
# Graph cars using blue points overlayed by a line 
plot(cars, type="o", col="blue")

# Create a title with a red, bold/italic font
title(main="Autos", col.main="red", font.main=4)
```


```{r linechart_3}
# Define 2 vectors
cars <- c(1, 3, 6, 4, 9)
trucks <- c(2, 5, 4, 5, 12)

# Graph cars using a y axis that ranges from 0 to 12
plot(cars, type="o", col="blue", ylim=c(0,12))

# Graph trucks with red dashed line and square points
lines(trucks, type="o", pch=22, lty=2, col="red")

# Create a title with a red, bold/italic font
title(main="Autos", col.main="red", font.main=4)
```


```{r linechart_4}
# Define 2 vectors
cars <- c(1, 3, 6, 4, 9)
trucks <- c(2, 5, 4, 5, 12)

# Calculate range from 0 to max value of cars and trucks
g_range <- range(0, cars, trucks)

# Graph autos using y axis that ranges from 0 to max 
# value in cars or trucks vector.  Turn off axes and 
# annotations (axis labels) so we can specify them ourself
plot(cars, type="o", col="blue", ylim=g_range, 
   axes=FALSE, ann=FALSE)

# Make x axis using Mon-Fri labels
axis(1, at=1:5, lab=c("Mon","Tue","Wed","Thu","Fri"))

# Make y axis with horizontal labels that display ticks at 
# every 4 marks. 4*0:g_range[2] is equivalent to c(0,4,8,12).
axis(2, las=1, at=4*0:g_range[2])

# Create box around plot
box()

# Graph trucks with red dashed line and square points
lines(trucks, type="o", pch=22, lty=2, col="red")

# Create a title with a red, bold/italic font
title(main="Autos", col.main="red", font.main=4)

# Label the x and y axes with dark green text
title(xlab="Days", col.lab=rgb(0,0.5,0))
title(ylab="Total", col.lab=rgb(0,0.5,0))

# Create a legend at (1, g_range[2]) that is slightly smaller 
# (cex) and uses the same line colors and points used by 
# the actual plots 
legend(1, g_range[2], c("cars","trucks"), cex=0.8, 
   col=c("blue","red"), pch=21:22, lty=1:2);
```


```{r barchart_1}
# Graph the cars
barplot(cars)
```

```{r barchart_2}
# Graph cars with specified labels for axes.  Use blue 
# borders and diagnal lines in bars.
barplot(autos_data$cars, main="Cars", xlab="Days",  
   ylab="Total", names.arg=c("Mon","Tue","Wed","Thu","Fri"), 
   border="blue", density=c(10,20,30,40,50))
```

```{r barchart_3}
# Graph autos with adjacent bars using rainbow colors
barplot(as.matrix(autos_data), main="Autos", ylab= "Total",
   beside=TRUE, col=rainbow(5))

# Place the legend at the top-left corner with no frame  
# using rainbow colors
legend("topleft", c("Mon","Tue","Wed","Thu","Fri"), cex=0.6, 
   bty="n", fill=rainbow(5));
```

```{r barchart_4}
# Expand right side of clipping rect to make room for the legend
par(xpd=T, mar=par()$mar+c(0,0,0,4))

# Graph autos (transposing the matrix) using heat colors,  
# put 10% of the space between each bar, and make labels  
# smaller with horizontal y-axis labels
barplot(t(autos_data), main="Autos", ylab="Total", 
   col=heat.colors(3), space=0.1, cex.axis=0.8, las=1,
   names.arg=c("Mon","Tue","Wed","Thu","Fri"), cex=0.8) 
   
# Place the legend at (6,30) using heat colors
legend(6, 30, names(autos_data), cex=0.8, fill=heat.colors(3));
   
# Restore default clipping rect
par(mar=c(5, 4, 4, 2) + 0.1)
```

```{r histogram_1}
# Create a histogram for suvs
hist(cars)
```

```{r histogram_2}
# Concatenate the three vectors
autos <- c(autos_data$cars, autos_data$trucks, 
   autos_data$suvs)

# Create a histogram for autos in light blue with the y axis
# ranging from 0-10
hist(autos, col="lightblue", ylim=c(0,10))
```


```{r histogram_3}
# Concatenate the three vectors
autos <- c(autos_data$cars, autos_data$trucks, 
   autos_data$suvs)

# Compute the largest y value used in the autos
max_num <- max(autos)

# Create a histogram for autos with fire colors, set breaks
# so each number is in its own group, make x axis range from
# 0-max_num, disable right-closing of cell intervals, set
# heading, and make y-axis labels horizontal
hist(autos, col=heat.colors(max_num), breaks=max_num, 
   xlim=c(0,max_num), right=F, main="Autos Histogram", las=1)
```

```{r histogram_4}
# Concatenate the three vectors
autos <- c(autos_data$cars, autos_data$trucks, 
   autos_data$suvs)

# Compute the largest y value used in the autos
max_num <- max(autos)

# Create uneven breaks
brk <- c(0,3,4,5,6,10,16)

# Create a histogram for autos with fire colors, set uneven
# breaks, make x axis range from 0-max_num, disable right-
# closing of cell intervals, set heading, make y-axis labels 
# horizontal, make axis labels smaller, make areas of each
# column proportional to the count
hist(autos, col=heat.colors(length(brk)), breaks=brk, 
   xlim=c(0,max_num), right=F, main="Probability Density", 
   las=1, cex.axis=0.8, freq=F)
```

```{r histogram_5}
r <- rlnorm(1000)

hist(r)
```

```{r histogram_6}
r <- rlnorm(1000)

# Get the distribution without plotting it using tighter breaks
h <- hist(r, plot=F, breaks=c(seq(0,max(r)+1, .1)))

# Plot the distribution using log scale on both axes, and use
# blue points
plot(h$counts, log="xy", pch=20, col="blue",
	main="Log-normal distribution",
	xlab="Value", ylab="Frequency")
```


```{r piechart_1}
# Create a pie chart for cars
pie(cars)
```

```{r piechart_2}
# Create a pie chart with defined heading and
# custom colors and labels
pie(cars, main="Cars", col=rainbow(length(cars)),
   labels=c("Mon","Tue","Wed","Thu","Fri"))
```

```{r piechart_3}
# Define some colors ideal for black & white print
colors <- c("white","grey70","grey90","grey50","black")

# Calculate the percentage for each day, rounded to one 
# decimal place
car_labels <- round(cars/sum(cars) * 100, 1)

# Concatenate a '%' char after each value
car_labels <- paste(car_labels, "%", sep="")

# Create a pie chart with defined heading and custom colors
# and labels
pie(cars, main="Cars", col=colors, labels=car_labels,
   cex=0.8)

# Create a legend at the right   
legend(1.5, 0.5, c("Mon","Tue","Wed","Thu","Fri"), cex=0.8, 
   fill=colors)
```

```{r dotchart_cars_1}
# Create a dotchart for autos
dotchart(t(autos_data))
```

```{r dotchart_cars_2}
# Create a colored dotchart for autos with smaller labels
dotchart(t(autos_data), color=c("red","blue","darkgreen"),
   main="Dotchart for Autos", cex=0.8)
```


```{r misc}
# Make an empty chart
plot(1, 1, xlim=c(1,5.5), ylim=c(0,7), type="n", ann=FALSE)

# Plot digits 0-4 with increasing size and color
text(1:5, rep(6,5), labels=c(0:4), cex=1:5, col=1:5)

# Plot symbols 0-4 with increasing size and color
points(1:5, rep(5,5), cex=1:5, col=1:5, pch=0:4)
text((1:5)+0.4, rep(5,5), cex=0.6, (0:4))

# Plot symbols 5-9 with labels
points(1:5, rep(4,5), cex=2, pch=(5:9))
text((1:5)+0.4, rep(4,5), cex=0.6, (5:9))

# Plot symbols 10-14 with labels
points(1:5, rep(3,5), cex=2, pch=(10:14))
text((1:5)+0.4, rep(3,5), cex=0.6, (10:14))

# Plot symbols 15-19 with labels
points(1:5, rep(2,5), cex=2, pch=(15:19))
text((1:5)+0.4, rep(2,5), cex=0.6, (15:19))

# Plot symbols 20-25 with labels
points((1:6)*0.8+0.2, rep(1,6), cex=2, pch=(20:25))
text((1:6)*0.8+0.5, rep(1,6), cex=0.6, (20:25))
```

```{r flowers_data_df}
flowers <- data.frame (treat  = c('tip','tip','tip','tip','tip','tip','tip','tip','tip','notip','notip','notip','notip','notip','notip','notip','notip','notip'),
                  nitrogen = c('medium','medium','medium','high','high','high','low','low','low','medium','medium','medium','high','high','high','low','low','low'),
                  block = c(1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1),
                  height = c(7.5,10.7,11.2,12.6,10,14.1,6.4,7.6,9.7,7.5,4.1,3.5,1.2,2.6,10.9,5.2,2.2,4.5),
                  weight = c(7.62,12.14,12.76,18.66,18.07,19.12,5.97,13.05,6.49,13.6,12.58,12.93,18.24,16.57,17.22,5.79,9.97,8.6),
                  leafarea = c(11.7,14.1,7.1,18.6,16.9,13.1,8.7,7.2,8.1,13.6,13.9,16.6,16.6,17.1,49.2,11,9.6,9.4),
                  shootarea = c(31.9,46,66.7,54,90.5,113.2,7.3,47.2,18,122.2,136.6,109.3,148.1,141.1,189.6,67.4,63.1,113.5),
                  flower = c(1,10,10,9,3,13,2,8,3,11,11,3,7,3,17,5,2,7)
                )

```


```{r scatterplots_1}
# Create a scatterplot for flowers weight
plot(flowers$weight)
```

```{r scatterplots_2}
# Create a scatterplot for flowers weight using thw with() function
with(flowers, plot(weight)) 
```

```{r scatterplots_3}
# Create a scatterplot for flowers weight and shootarea
plot(x = flowers$weight, y = flowers$shootarea)
```

```{r scatterplots_4}
#reference: https://intro2r.com/custom_plot.html
# Create a scatterplot for flowers weight and shootarea with legend
par(mar = c(4.1, 4.4, 4.1, 1.9), xaxs = "i", yaxs = "i")
plot(flowers$weight, flowers$shootarea,
       type = "n",
       xlab = "weight (g)",
       ylab = expression(paste("shoot area (cm"^"2",")")),
       xlim = c(0, 30), ylim = c(0, 200), bty = "l",
       las = 1, cex.axis = 0.8, tcl = -0.2)
points(x = flowers$weight[flowers$nitrogen == "low"],
       y = flowers$shootarea[flowers$nitrogen == "low"],
       pch = 16, col = "deepskyblue")

points(x = flowers$weight[flowers$nitrogen == "medium"],
       y = flowers$shootarea[flowers$nitrogen == "medium"],
       pch = 16, col = "yellowgreen")
points(x = flowers$weight[flowers$nitrogen == "high"],
       y = flowers$shootarea[flowers$nitrogen == "high"],
       pch = 16, col = "deeppink3")
text(x = 28, y = 190, label = "A", cex = 2)

leg_cols <- c("deepskyblue", "yellowgreen", "deeppink3")
leg_sym <- c(16, 16, 16)
leg_lab <- c("low", "medium", "high")

legend(x = 1, y = 200, col = leg_cols, pch = leg_sym, 
        legend = leg_lab, bty = "n", 
        title = "Nitrogen level")
```



```{r linegraph_two_vector}
# Plot two vectors of numbers using different type = values
my_x <- 1:10
my_y <- seq(from = 1, to = 20, by = 2)

par(mfrow = c(2, 2))
plot(my_x, my_y, type = "l")
plot(my_x, my_y, type = "b")
plot(my_x, my_y, type = "o")
plot(my_x, my_y, type = "c")
```


```{r kernel_density_curve}
# add a kernel density curve to the plot.
dens <- density(flowers$height)
hist(flowers$height, breaks = brk, main = "petunia height",
      freq = FALSE)
lines(dens)
```
Box Plot
Reference: http://www.sthda.com/english/wiki/ggplot2-box-plot-quick-start-guide-r-software-and-data-visualization

```{r boxplot_1}
#  Create a boxplot for flowers weight
boxplot(flowers$weight, ylab = "weight (g)",
        col = c("white"),
         border = c("#E69F00"))
```

```{r boxplot_2}
#  Create a boxplot to show distribution of a variable changes between different levels of a factor
boxplot(weight ~ nitrogen, data = flowers, 
        ylab = "weight (g)", xlab = "nitrogen level",
        col = c("#999999", "#E69F00", "#56B4E9"),
        border = c("black"))

```

```{r boxplot_3}
# boxplot with our factor levels going from low to high.
flowers$nitrogen <- factor(flowers$nitrogen, 
                            levels = c("low", "medium", "high"))
boxplot(weight ~ nitrogen, data = flowers, 
          ylab = "weight (g)", xlab = "nitrogen level")
```

```{r boxplot_4}
# group variables by two factors in the same plot.
boxplot(weight ~ nitrogen * treat, data = flowers, 
         ylab = "weight (g)", xlab = "nitrogen level")
```


```{r boxplot_5}
# reduce the font size of the tick mark labels in the plot
boxplot(weight ~ nitrogen * treat, data = flowers, 
         ylab = "weight (g)", xlab = "nitrogen level", 
         cex.axis = 0.7)
```

```{r violin_plot_1}
vioplot(weight ~ nitrogen, data = flowers, 
         ylab = "weight (g)", xlab = "nitrogen level",
         col = "lightblue")
```

```{r dotchart_flowers_1}
#  Create a dotchart for flowers height
dotchart(flowers$height)
```

```{r dotchart_flowers_2}
#using the groups = argument to group variables value
dotchart(flowers$height, groups = flowers$nitrogen)
```

The pairs() function creates a multi-panel scatterplot (sometimes called a scatterplot matrix) which plots all combinations of variables.

```{r pairsplot_1}
#multi-panel scatterplot of all of the numeric variables in our flowers data frame

pairs(flowers[, c("height", "weight", "leafarea", 
                "shootarea", "flower")])
```

```{r pairsplot_2}
#add a LOWESS (locally weighted scatterplot smoothing) smoother to each of the panels we just need to add the argument

pairs(flowers[, c("height", "weight", "leafarea", 
                "shootarea", "flower")], 
                 panel = panel.smooth)
```


conditioning plot (also known as conditional scatterplot plot) 
```{r coplots_1}
#two variables but each plot is conditioned (|) by a third variable. 

coplot(flower ~ weight|leafarea, data = flowers)
```


```{r coplots_2}
#use two conditioning variables (either numeric or factors)

coplot(flower ~ weight|nitrogen * treat, data = flowers)
```

```{r coplots_3}
#apply functions to each of the panels using the panel = argument.

coplot(flower ~ weight|nitrogen * treat, data = flowers,
        panel = function(x, y, ...) {
        points(x, y, ...)
        abline(lm(y ~ x), col = "blue")})
```

Lattice plots
```{r lattice_1}
#recreate the frequency histogram of the height variable in our flowers data frame using the histogram() function.

histogram(~ height, type = "count", data = flowers)
```

```{r lattice_2}
#boxplot of our weight variable for each level of nitrogen using the bwplot() function.

bwplot(weight ~ nitrogen, data = flowers)
```

```{r lattice_3}
# plot graphs in multiple panels

histogram(~ height | nitrogen, type = "count", data = flowers)
```

```{r lattice_4}
#change the layout of our plots we can use the layout = argument.

histogram(~ height | nitrogen, type = "count", 
           layout = c(1, 3), data = flowers)
```

```{r lattice_5}
#create conditional boxplots using the same logic.

bwplot(weight ~ nitrogen | block, data = flowers)
```

```{r lattice_6}
#create a scatter plot of height against weight for each level of nitrogen and treat

xyplot(height ~ weight | nitrogen * treat, data = flowers)
```


Multiple graphs

```{r multigraphs_1}
#1 row and 2 columns.

par(mfrow = c(1, 2))
plot(flowers$weight, flowers$shootarea, xlab = "weight",
      ylab = "shoot area")
boxplot(shootarea ~ nitrogen, data = flowers, cex.axis = 0.6)
```

```{r multigraphs_2}
#2 row and 2 columns.

par(mfrow = c(2, 2))
plot(flowers$weight, flowers$shootarea, xlab = "weight",
      ylab = "shoot area")
boxplot(shootarea ~ nitrogen, cex.axis = 0.8, data = flowers)
hist(flowers$weight, main ="")
dotchart(flowers$weight)
```


The layout() function 
Allows you to split your plotting device up into different sized regions and can be used to build complex figures. 
```{r multigraphs_3}
#Set the graph layout
layout_mat <- matrix(c(2, 0, 1, 3), nrow = 2, ncol = 2,
                      byrow = TRUE)

my_lay <- layout(mat = layout_mat, 
                 heights = c(1, 3),
                 widths = c(3, 1), respect =TRUE)

#create our three plots and change the figure margins for each of the figures using the par(mar = )

par(mar = c(4, 4, 0, 0))
plot(flowers$weight, flowers$shootarea, 
     xlab = "weight (g)", ylab = "shoot area (cm2)")
par(mar = c(0, 4, 0, 0))
boxplot(flowers$weight, horizontal = TRUE, frame = FALSE,
        axes =FALSE)
par(mar = c(4, 0, 0, 0))
boxplot(flowers$shootarea, frame = FALSE, axes = FALSE)
```


GGPLOT

```{r ggplot_1}
# Including aesthetics for x and y axes as well as specifying the dataset
ggplot(aes(x = weight, y = shootarea), data = flowers) +
        geom_point() +    # Adding a geom to display data as point data
        geom_line()     # Adding geom_line


```

```{r ggplot_2}
ggplot(aes(x = weight, y = shootarea), data = flowers) +
  geom_point() +
    geom_smooth()    # Changing to geom_smooth
```

```{r ggplot_3}
ggplot(aes(x = weight, y = shootarea), data = flowers) +
  geom_point() +
  # Including colour argument in aes()
  geom_smooth(aes(colour = nitrogen), method = "lm", se = FALSE)
```


```{r ggplot_4}
# Moved colour = nitrogen into the universal ggplot()
ggplot(aes(x = weight, y = shootarea, colour = nitrogen), data = flowers) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```


```{r ggplot_5}
ggplot(aes(x = weight, y = shootarea, colour = nitrogen), data = flowers) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
    # Splitting the single figure into multiple depending on treatment
  facet_wrap(~ treat)
```

```{r ggplot_6}
ggplot(aes(x = weight, y = shootarea, colour = nitrogen), data = flowers) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
    # Adding "block" to formula
  facet_wrap(~ treat + block)


```


Reference: 
Be Awesome in ggplot2: A Practical Guide to be Highly Effective - R software and data visualization
Link: http://www.sthda.com/english/wiki/be-awesome-in-ggplot2-a-practical-guide-to-be-highly-effective-r-software-and-data-visualization
